---
title: "simulation_plots"
author: ""
date: ""
output: html_document
---

```{r}
root <- has_file("README.md")$make_fix_file()
source(root("spc","evaluation.R")) 
```

# Poisson Model


- Compare different SPC proportions for single SPC

```{r}
load("poisson_model.rda")
updated_data <- updated_data %>% filter(method %in% c("single 0.1-SPC", "single 0.3-SPC", "single 0.5-SPC", "single 0.7-SPC", "single 0.9-SPC"))
eval_checks(updated_data, discr_name = "emp_mean",  model_name = "Poisson model")
eval_checks(updated_data, discr_name = "sec_moment",  model_name = "Poisson model")
eval_checks(updated_data, discr_name = "third_moment",  model_name = "Poisson model")
eval_checks(updated_data, discr_name = "mse",  model_name = "Poisson model")
```


- Compare different SPC proportions for divided SPC with $k = N^{0.49}$.
```{r}
load("poisson_model.rda")
updated_data <- updated_data %>% filter(method %in% c("divided 0.1-SPC", "divided 0.3-SPC", "divided 0.5-SPC", "divided 0.7-SPC", "divided 0.9-SPC"))

eval_checks(updated_data, discr_name = "emp_mean", model_name = "Poisson model")
eval_checks(updated_data, discr_name = "sec_moment",model_name = "Poisson model")
eval_checks(updated_data, discr_name = "third_moment",model_name = "Poisson model")
eval_checks(updated_data, discr_name = "mse", model_name = "Poisson model")

```


- Compare different k's for divided SPC with SPC proportion $q = 0.5$

```{r}
load("poisson_model.rda")
updated_data <- updated_data %>% filter(method %in% c("divided 0.5-SPC, k = N^(.4)", "divided 0.5-SPC, k = N^(.49)", "divided 0.5-SPC, k = N^(.6)", "divided 0.5-SPC, k = N^(.8)"))
eval_checks(updated_data, discr_name = "emp_mean", model_name = "Poisson model")
eval_checks(updated_data, discr_name = "sec_moment",model_name = "Poisson model")
eval_checks(updated_data, discr_name = "third_moment",model_name = "Poisson model")
eval_checks(updated_data, discr_name = "mse", model_name = "Poisson model")
```


- Compare `PPC`, `single 0.5-SPC`, `divided 0.5-SPC` power/ test size
```{r}
load("poisson_model.rda")
updated_data <- updated_data %>% filter(method %in% c("PPC", "single 0.5-SPC", "divided 0.5-SPC"))
eval_checks(updated_data, discr_name = "emp_mean", model_name = "Poisson model")
eval_checks(updated_data, discr_name = "sec_moment",model_name = "Poisson model")
eval_checks(updated_data, discr_name = "third_moment",model_name = "Poisson model")
eval_checks(updated_data, discr_name = "mse", model_name = "Poisson model")

qq_plot_comparison(updated_data, discr_name = "emp_mean")
qq_plot_comparison(updated_data, discr_name = "sec_moment")
qq_plot_comparison(updated_data, discr_name = "third_moment")
qq_plot_comparison(updated_data, discr_name = "mse")
```


```{r}
load("poisson_mismatch_comparison.rda")
eval_checks_clean_mismatch(updated_data, discr_name = "emp_mean")
eval_checks_clean_mismatch(updated_data, discr_name = "sec_moment")

```

- Compare `PPC`, `single 0.9-SPC`, `divided 0.9-SPC` adjusted priorESS 
```{r}
load("poisson_model_ess.rda")
updated_data <- updated_data %>% filter(true_theta == 25, quantile == 0.95)
plot_ts_ESS(updated_data, model_name = "Poisson", discr_fun_name = "emp_mean") 
plot_ts_ESS(updated_data, model_name = "Poisson", discr_fun_name = "sec_moment") 
```




# Gaussian Model

- Compare different SPC proportions for single SPC
```{r}
load("gaussian_model.rda")
updated_data <- updated_data %>% filter(method %in% c("single 0.1-SPC", "single 0.3-SPC", "single 0.5-SPC", "single 0.7-SPC", "single 0.9-SPC"))
eval_checks(updated_data, discr_name = "emp_mean",  model_name = "Gaussian model")
eval_checks(updated_data, discr_name = "sec_moment",  model_name = "Gaussian model")
eval_checks(updated_data, discr_name = "quantile_0.75",  model_name = "Gaussian model")
eval_checks(updated_data, discr_name = "mse",  model_name = "Gaussian model")
```

- Compare different SPC proportions for divided SPC with $k = N^{0.49}$.
```{r}
load("gaussian_model.rda")
updated_data <- updated_data %>% filter(method %in% c("divided 0.1-SPC", "divided 0.3-SPC", "divided 0.5-SPC", "divided 0.7-SPC", "divided 0.9-SPC"))
eval_checks(updated_data, discr_name = "emp_mean",  model_name = "Gaussian model")
eval_checks(updated_data, discr_name = "sec_moment",  model_name = "Gaussian model")
eval_checks(updated_data, discr_name = "quantile_0.75",  model_name = "Gaussian model")
eval_checks(updated_data, discr_name = "mse",  model_name = "Gaussian model")

```


- Compare different k's for divided SPC with SPC proportion $q = 0.9$

```{r}
load("gaussian_model.rda")
updated_data <- updated_data %>% filter(method %in% c("divided 0.5-SPC, k = N^(.4)", "divided 0.5-SPC, k = N^(.49)", "divided 0.5-SPC, k = N^(.6)", "divided 0.5-SPC, k = N^(.8)"))
eval_checks(updated_data, discr_name = "quantile_0.75",  model_name = "Gaussian model")
eval_checks(updated_data, discr_name = "sec_moment",  model_name = "Gaussian model")
eval_checks(updated_data, discr_name = "emp_mean",  model_name = "Gaussian model")
eval_checks(updated_data, discr_name = "mse",  model_name = "Gaussian model")
```


- Compare `PPC`, `single 0.9-SPC`, `divided 0.9-SPC` power/ test size
```{r}
load("gaussian_model.rda")
updated_data <- updated_data %>% filter(method %in% c("PPC", "single 0.5-SPC", "divided 0.5-SPC"))
eval_checks(updated_data, discr_name = "emp_mean",  model_name = "Gaussian model")
eval_checks(updated_data, discr_name = "quantile_0.75",  model_name = "Gaussian model")
eval_checks(updated_data, discr_name = "sec_moment",  model_name = "Gaussian model")
eval_checks(updated_data, discr_name = "mse",  model_name = "Gaussian model")

qq_plot_comparison(updated_data, discr_name = "emp_mean")
qq_plot_comparison(updated_data, discr_name = "quantile_0.75")
qq_plot_comparison(updated_data, discr_name = "sec_moment")
qq_plot_comparison(updated_data, discr_name = "mse")
```


# Gaussian Hierachical Model

### Scenario 1: well-specified model
```{r well-specified}
load("hier_well_specified.rda")
updated_data <- updated_data %>% filter(num_of_indi == 8)
size_plot_hier_fixJ_clean(updated_data, discr_name = "grand_mean", iter = 200)
size_plot_hier_fixJ_clean(updated_data, discr_name = "q0.75_group_means", iter = 200)
size_plot_hier_fixJ_clean(updated_data, discr_name = "mean_group_q75", iter = 200)

qq_plot_comparison_hier(updated_data, fixed_num_of_indi = 8,  discr_name = "grand_mean", iter = 200)
qq_plot_comparison_hier(updated_data, fixed_num_of_indi = 8,  discr_name = "q0.75_group_means", iter = 200)
qq_plot_comparison_hier(updated_data, fixed_num_of_indi = 8,  discr_name = "mean_group_q75", iter = 200)
```
### Scenario 2: Misspecified across groups
```{r S1-hier_cr, message = FALSE}
load("hier_cr_jeff.rda")
updated_data <- updated_data %>% filter(num_of_indi == 8)
eval_checks_hier_fixJ_clean(updated_data, discr_name = "", iter = 200)

load("hier_cr_jeff.rda")
updated_data <- updated_data %>% filter(num_of_groups == 20)
eval_checks_hier_fixI_clean(updated_data, discr_name = "", iter = 200)

```
### Scenario 3: Misspecified within groups
```{r S1-hier_cr, message = FALSE}
load("hier_wi_gr_var.rda")
updated_data <- updated_data %>% filter(num_of_indi == 8)
eval_checks_hier_fixJ_clean(updated_data, discr_name = "", iter = 200)

load("hier_wi_gr_var.rda")
updated_data <- updated_data %>% filter(num_of_groups == 20)
eval_checks_hier_fixI_clean(updated_data, discr_name = "", iter = 200)

```


### Scenario 4: Misspecified both across and within groups
```{r S1-hier_cr, message = FALSE}
load("hier_wi_lognorm.rda")
updated_data <- updated_data %>% filter(num_of_indi == 8)
eval_checks_hier_fixJ_clean(updated_data, discr_name = "", iter = 200)

load("hier_wi_lognorm.rda")
updated_data <- updated_data %>% filter(num_of_groups == 20)
eval_checks_hier_fixI_clean(updated_data, discr_name = "", iter = 200)

```
